#Practical Machine Learning Final Project

## Synopsis:

Accelerometer data from participants performing barbell lifts five different ways was collected to 
create a model that was used to predict the classe of barbell lift for twenty sets of acceleromter data.

## Data Cleaning:

A training set and testing set were obtained from http://groupware.les.inf.puc-rio.br/har 
under the Weight Lifting Exercise data set.

The training set is from http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The testing set of twenty sets of data for which a classe is to be predicted is from
http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

Before the prediction can be made, the training set is divided into a smaller training set (75%)
and validation set (25%)

```{r echo=TRUE}
setwd("C:/Users/mbonifac/Desktop/Coursera/R_Portfolio")
trainmaster<-read.csv("./pml-training.csv")
#Remove the columns that are blank or have NA
working_train<-trainmaster[lapply(trainmaster,function(x) sum(is.na(x))/length(x))<0.1]
working_train[working_train == ""] <- NA
working_train<-working_train[lapply(working_train,function(x) sum(is.na(x))/length(x))<0.1]

#Remove columns that are known to not affect classe, (row number, user_name, raw_timestamp_part_1, 
#raw_timestamp_part_2, cvtd_timestamp, new_window, num_window)
working_train<-working_train[,-c(1:7)]

#Divide data into a training set and a validation set (75% training/25% validation)
library(caret)
library(kernlab)
set.seed(125)
inTrain<-createDataPartition(y=working_train$classe, p=0.75, list=FALSE)
train_train<-working_train[inTrain,]
train_validate<-working_train[-inTrain,]
```

The testing set is prepared in the same manner as the working_train set with unnecessary columns removed.

```{r echo=TRUE}
testmaster<-read.csv("./pml-testing.csv")

#Remove the columns that are blank or have NA
working_test<-testmaster[lapply(testmaster,function(x) sum(is.na(x))/length(x))<0.1]
working_test[working_test == ""] <- NA
working_test<-working_test[lapply(working_test,function(x) sum(is.na(x))/length(x))<0.1]

#Remove columns that are known to not affect classe, (row number, user_name, raw_timestamp_part_1, 
#raw_timestamp_part_2, cvtd_timestamp, new_window, num_window)
working_test<-working_test[,-c(1:7)]
```

## Determination of Model:

The first model to attempt is the decision tree.

```{r echo=TRUE}
model_tree<-train(classe~.,method="rpart",data=train_train)
library(rattle)
fancyRpartPlot(model_tree$finalModel)
```

The decision tree model does not appear to identify classe D.  Predict using the train_validate set to confirm.

```{r echo=TRUE}
valid_tree<-predict(model_tree,newdata=train_validate)
confusionMatrix(valid_tree,train_validate$classe)
```

The accuracy of the decision tree model is low.  The boosted model is then tried.

```{r echo=TRUE}
model_boost<-train(classe~.,method="gbm",data=train_train, verbose=FALSE)
valid_boost<-predict(model_boost,newdata=train_validate)
confusionMatrix(valid_boost,train_validate$classe)
```

## Prediction

With a much higher accuracy, the boosted model is considered for prediction of the working_test set.
The boosted model is used to predict classe for the data in working_test.

```{r echo=TRUE}
test_prediction<-predict(model_boost,newdata=working_test)
Prediction_Results<-data.frame(prediction=test_prediction)
Prediction_Results
```
